---
title: "Movie Recommendation Group Project"
---


```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(readxl)
library(ggpubr)

# Import data
raw_links <- read.csv("C:/NEW/MAST 6251/the-movies-dataset/links.csv")
raw_movies_metadata <- read.csv("C:/NEW/MAST 6251/the-movies-dataset/movies_metadata1.csv")
raw_ratings <- read.csv("C:/NEW/MAST 6251/the-movies-dataset/ratings.csv")

```

## Introduction
The purpose of the project was to predict a movie rating score that would determine whether a movie is considered a good movie or not. The targeted user was online audio/video platforms such as Netflix, Amazon Prime Video, etc.
A logistic regression model was made after testing the 45000 movie rating found in the Full MovieLens Dataset consisting of movies released on or before July 2017.
The model includes movie popularity, runtime, released year, budget and production countries as influence factors. The effect of each variable to the movie rating score (good score) is explained in detail below along with a series of new findings.

```{r include=FALSE}
# remove duplicate and remove meaningless data of metadata (all meaningless rows happen in the same row for other columns)
raw_movies_metadata  = raw_movies_metadata[!duplicated(raw_movies_metadata$id), ]
```

```{r include=FALSE}
# use tmbdid to join data
# remove duplicate and null value from links
raw_links  = raw_links[!duplicated(raw_links$tmdbId)& !is.na(raw_links$tmdbId),]
```

```{r include=FALSE}
# aggregate average of rating by movie id
avg_rating = raw_ratings %>% group_by(movieId) %>% summarize(avg_rating = mean(rating, na.rm = TRUE))
```

```{r include=FALSE}
# join three tables together (leftjoin)
dataset = left_join(raw_movies_metadata, raw_links, by=c('id'='tmdbId'))
dataset = left_join(dataset, avg_rating, by=c('movieId'='movieId'))
```

```{r include=FALSE}
# remove ratings with null after leftjoin (result in innerjoin)
dataset = dataset[!is.na(dataset$avg_rating),]
```

```{r include=FALSE}
# binary variable classifications
dataset1=mutate(dataset, belongs_to_collection = ifelse(belongs_to_collection=="",0,1))
dataset1=mutate(dataset1, original_language=ifelse(original_language=='en','English','Non English'))
dataset1=mutate(dataset1, production_companies= ifelse(production_companies=="",0,1))
dataset1=mutate(dataset1,adult=ifelse(adult==TRUE,1,0))
dataset1=mutate(dataset1,video=ifelse(video==FALSE,0,1))
```

```{r include=FALSE}
# add columns for month and year
dataset1$release_year=substr(as.Date(dataset1$release_date,"%m/%d/%Y"),1,4)
dataset1$release_month=substr(as.Date(dataset1$release_date,"%m/%d/%Y"),6,7)
```

```{r include=FALSE}
# binary variable agian
dataset1=mutate(dataset1, status= ifelse(status=="Released","Released","Not Released"))
```

```{r include=FALSE}
# production country with US is defined as "include US"
dataset1=mutate(dataset1, production_countries= ifelse(grepl('US', as.character((dataset1$production_countries)))=='TRUE','Include US','Non US'))
nrow(dataset1)
```

```{r include=FALSE}
# add profit(revenue minus budget)
dataset1$profit=dataset1$revenue-dataset1$budget
```

```{r include=FALSE}
# drop little business value variables
dataset1=subset(dataset1,select=-c(homepage,overview,original_title,poster_path,imdb_id,tagline,id,movieId))
```

```{r include=FALSE}
# set variables to character
dataset1$adult<-as.character(dataset1$adult)
dataset1$belongs_to_collection<-as.character(dataset1$belongs_to_collection)
dataset1$production_companies<-as.character(dataset1$production_companies)
dataset1$video<-as.character(dataset1$video)

```

```{r include=FALSE}
#detect and remove outliers
outvars <- unlist(lapply(dataset1, is.numeric))
dataset1[outvars] <- data.frame(lapply(dataset1[outvars], function(x){
  ifelse((x < 0) | x > (mean(x,na.rm=TRUE) + 3*sd(x,na.rm=TRUE)), NA, x)
}))
dataset1= na.omit(dataset1)
str(dataset1)
```

```{r include=FALSE}
# set variables to numeric
dataset1$release_year<-as.numeric(dataset1$release_year)
dataset1$adult<-as.numeric(dataset1$adult)
dataset1$belongs_to_collection<-as.numeric(dataset1$belongs_to_collection)
dataset1$video<-as.numeric(dataset1$video)
dataset1$production_companies<-as.numeric(dataset1$production_companies)
```

## Predictive Model Approach

Budget - The budget in which the movie was made
I can see that the budget of movies will influence the average rating, therefore influencing the decision that whether they are good movies or not. However, the result shows that movies’ budgets may have a negative effect on movies’ ratings. If the filmmakers invest 1 dollar more in the movie, there will be 1.04 decrease in a good score. In other words, the more budget a movie has, the more likely that the probability of the good movie decreases.
Maybe the “artistic cheapening” phenomenon accounts for this situation. The blockbuster may seem to use CGI technology to make up for a lack of plot and dialogue, leading to a decrease in the movies’ quality.
Therefore, because the average of movies’ budget is 1047180 dollars, if a movie’s budget is higher than approximately 1000,000 dollars, maybe there will be an increased likelihood that this movie is not a good movie.

Although we successfully built the model, from the summary of GLM regression, I find that popularity,runtime and budget leave a negative influence on the good score, which is little inconsistent with common business meaning. However, the p-value for these variables is significant. AIC also is relatively smallest among other models. Residual plot shows little fitted discernible pattern, too. Therefore, from the class of endogeneity, I guess that our model may have some omit ted variables, which may have left some influence on dependent variables and good core. In order to solve this problem, we may need to set up an instrumental variable such as the number of actors, which is unlikely to directly influence the goodScore. But, more actors acted in movies may be more likely to increase the budget (correlated with budget, x) and use two-stage least squares to get an unbiased estimate of model.

```{r}
# create variable goodScore
dataset1 = dataset1 %>% mutate(goodScore = (avg_rating >= 4)*1)
```


```{r}
# linear model
linReg1 = lm(avg_rating~adult+ belongs_to_collection  + original_language + popularity + production_companies+runtime+vote_average+release_year+profit,dataset1)
plot(linReg1,which = 1)
summary(linReg1)

# log y model
log = lm(log(avg_rating) ~ adult+ belongs_to_collection  + original_language + popularity + production_companies+runtime+vote_average+release_year+profit,dataset1)
summary(log)
plot(log,which=1)

# glm first try
logReg1 = glm(goodScore ~ adult+ belongs_to_collection  + original_language + popularity + production_companies+runtime+vote_average+release_year+profit,dataset1,family=binomial)
summary(logReg1)
plot(logReg1,which=1)

# glm second try
logReg2 = glm(goodScore ~  belongs_to_collection  + original_language + popularity +I(popularity^2)+runtime+vote_average+release_year+profit,dataset1,family=binomial)
summary(logReg2)
plot(logReg2,which=1)

# glm third try (Final decision)
logReg3 = glm(goodScore ~  popularity +runtime + release_year + budget + production_countries,dataset1,family=binomial)
summary(logReg3)
plot(logReg3,which=1)
```

```{r}
# figure out the model accuracy

# Psuedo R-squared
logRegFin = logReg3 
1 - logRegFin$deviance/logRegFin$null.deviance

# Classification Matrix
pred = predict(logRegFin,type="response") 
pred[pred>=.5] = "goodScore"
pred[pred!="goodScore"] = "badScore"
classMatrix = table(pred,dataset1$goodScore) 
classMatrix

# good/bad prediction accuracy rate
sum(diag(classMatrix))/sum(classMatrix)
```

## Use Cases

After observing and predicting history movie data, we recommend that the users should choose the movies that have last 10% popularity, last 10% budget, last 10% runtime, top 10% release year and are produced in non-us countries considering the negative and positive effects on showing probability of good movie.  
In other words, if the users see a non-us movie that has popularity lower than 0.078, budget equals to 0, runtime less than 62 and is later than 2014, there is relatively large 12% probability that this movie is a good movie and has the high average rating greater than 4.



